Author disambiguation of large-scale semi-labeled collections of scientific articles

### 2015-08-17 paper.tex
remove mentions of matching

Author name disambiguation in bibliographic databases is the problem of
grouping together scientific publications written by a same person, accounting
for potential homonyms and/or synonyms. Among solutions to this problem,
digital libraries are increasingly offering tools for authors to manually
curate their publications and claim which ones are theirs. Indirectly, these
tools allow for the inexpensive collection of large annotated training data,
which can be further leveraged to build a complementary automated disambiguation system
capable of inferring patterns for identifying publications written by a same
person.  Building upon more than 1 million crowdsourced annotations, we
propose an automated author disambiguation solution exploiting this data (i) to learn
an accurate classifier for identifying corefering authors, (ii) to guide the
clustering of scientific publications by distinct authors in a semi-supervised
way and finally (iii) to match the resulting clusters with existing profiles.
To the best of our knowledge, our analysis is the first to be carried out on
data of this size and coverage. It reveals that... \glnote{To be completed with
actual results.}

@@

\glnote{Note sure if we should talk about matching in this paper?}
\glnote{Insist on the complementarity of having both crowdsourced and automated disambiguation.}

@@

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or by developing services inviting authors to register
their publications themselves (e.g., Google Scholar, Inspire-HEP, ...).
Recently, efforts also started to create persistent digital identifiers
assigned to researchers (e.g., ORCID, ResearchedID, ...), with the further
objective to embed these identifiers in the submission workflow of publishers
or repositeries (e.g., Elsevier, arXiv, Inspire-HEP, ...), thereby univocally
solving any disambiguation issue. With the large cost of centralized manual
authorship control, or until the larger adoption of crowdsourced solutions, the
impact of these efforts are unfortunately limited by the efficiency, motivation
and integrity of their active contributors. Similarly, the success of
persistent digital identifier efforts is conditioned to a large and ubiquituous
adoption by both researchers and publishers. For these reasons, fully automated
machine learning-based methods have been proposed during the past decade to
provide immediate, less costly and efficient solutions to author
disambiguation. In this work, our goal is to explore and demonstrate how both
approaches can coexist and benefit from each other.  In particular, we study
how labeled data obtained through manual curation (either centralized or
crowdsourced) can be exploited (i) to learn an accurate classifier for
identifying corefering authors, (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way and finally (iii) to
match the resulting clusters with existing profiles. Our analysis of parameters
and features on this large training set reveals that... \glnote{To be completed
with actual results.} \glnote{Also mention error analysis, suggesting that
uncertain disambiguation results should be suggested to human curators, thereby
closing the loop and creating a vertuous circle.}

@@

In this work, author disambiguation is cast into a semi-supervised clustering
problem, as inspired from several previous works described in Section~\ref
{related-works}. Our algorithm is made of four elements: (i) a blocking
function $b$ whose goal is to rhoughly pre-cluster signatures ${\cal S}$ into smaller groups in order to
reduce computational complexity; (ii) the construction of a similarity function
$d$ between signatures using supervised learning; (iii) the
semi-supervised clustering of all signatures within a same block, using $d$; (iv) the matching
of predicted clusters to existing profiles in a database.

@@

\subsection{Matching}
% general description in a real system

>>>>

Author name disambiguation in bibliographic databases is the problem of
grouping together scientific publications written by a same person, accounting
for potential homonyms and/or synonyms. Among solutions to this problem,
digital libraries are increasingly offering tools for authors to manually
curate their publications and claim which ones are theirs. Indirectly, these
tools allow for the inexpensive collection of large annotated training data,
which can be further leveraged to build a complementary automated
disambiguation system capable of inferring patterns for identifying
publications written by a same person.  Building upon more than 1 million
crowdsourced annotations, we propose an automated author disambiguation
solution exploiting this data (i) to learn an accurate classifier for
identifying corefering authors and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way. To the best of our
knowledge, our analysis is the first to be carried out on data of this size and
coverage. It reveals that... \glnote{To be completed with actual results.}

@@

\glnote{Insist on the complementarity of having both crowdsourced and automated disambiguation.}

@@

Efforts and solutions to author disambiguation have been proposed from various
communities \citep{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring collaborators or by developing services inviting authors to register
their publications themselves (e.g., Google Scholar, Inspire-HEP, ...).
Recently, efforts also started to create persistent digital identifiers
assigned to researchers (e.g., ORCID, ResearchedID, ...), with the further
objective to embed these identifiers in the submission workflow of publishers
or repositeries (e.g., Elsevier, arXiv, Inspire-HEP, ...), thereby univocally
solving any disambiguation issue. With the large cost of centralized manual
authorship control, or until the larger adoption of crowdsourced solutions, the
impact of these efforts are unfortunately limited by the efficiency, motivation
and integrity of their active contributors. Similarly, the success of
persistent digital identifier efforts is conditioned to a large and ubiquituous
adoption by both researchers and publishers. For these reasons, fully automated
machine learning-based methods have been proposed during the past decade to
provide immediate, less costly and efficient solutions to author
disambiguation. In this work, our goal is to explore and demonstrate how both
approaches can coexist and benefit from each other.  In particular, we study
how labeled data obtained through manual curation (either centralized or
crowdsourced) can be exploited (i) to learn an accurate classifier for
identifying corefering authors and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way. Our analysis of parameters
and features on this large training set reveals that... \glnote{To be completed
with actual results.} \glnote{Also mention error analysis, suggesting that
uncertain disambiguation results should be suggested to human curators, thereby
closing the loop and creating a vertuous circle.}

@@

In this work, author disambiguation is cast into a semi-supervised clustering
problem, as inspired from several previous works described in Section~\ref
{related-works}. Our algorithm is made of three elements: (i) a blocking
function $b$ whose goal is to rhoughly pre-cluster signatures ${\cal S}$ into smaller groups in order to
reduce computational complexity; (ii) the construction of a similarity function
$d$ between signatures using supervised learning; (iii) the
semi-supervised clustering of all signatures within a same block, using $d$ as a pseudo distance metric.

@@

% \subsection{Matching}
% general description in a real system
% GL: out of scope, let us mention that in Conclusions instead
